NumPy---Array Computation for Python

### 2020-01-05 body.tex
Minor fix
Update section title
Update outline

\section{Introduction}

@@

\subsection{History}

@@

\subsection{Enabling factors}

@@

\subsection{Hybrid development}

@@

% BIDS -- UCB
% https://www.moore.org/grant-detail?grantId=GBMF5447
% $645,020 in 2016
% https://sloan.org/grant-detail/8222
% $659,359 in 2017
% https://bids.berkeley.edu/news/bids-receives-sloan-foundation-grant-contribute-numpy-development
% http://doi.org/10.5281/zenodo.3585761
% http://doi.org/10.5281/zenodo.3585767
In 2017, NumPy received its first large grants totaling 1.3M USD from the
Gordon \& Betty Moore and the Alfred P. Sloan foundations.
These two grants focus on addressing the technical debt accrued over the years and
on setting in place standards and architecture to encourage more sustainable development.
% CZI -- NumFOCUS/QuanSight
% https://chanzuckerberg.com/eoss/proposals/strengthening-numpys-foundations-growing-beyond-code/
% NumPy and OpenBLAS received $195,000 in 2019
% https://labs.quansight.org/blog/2019/11/numpy-openblas-CZI-grant/
NumPy received a third grant for 200K USD from the Chan Zuckerberg
Initiative at the end of 2019.
This grant focuses on better serving NumPy's large number of beginning
to intermediate level users and on growing the community of NumPy
contributors.
Finally, the project receives around 10K USD annually from Tidelift, which is
typically used to fund documentation writers and website developers.

@@

\section{The \code{ndarray} data structure}

@@

The \code{ndarray} data structure stores regularly spaced elements of a single
data type in a contiguous block of memory; along with meta-data such as shape,
it allows for the efficient representation of $n$-dimensional arrays.
More details about the \code{ndarray} data structure are given in ``The NumPy array:
a structure for efficient numerical computation''\cite{vanderwalt2011numpy}
and our online documentation.

The need for such a data structure arises often in science.
For instance, they may store measurements from an experiment or simulation
at discrete time points.
Data analysis frequently involves operations $(+, -, \times, /, \log, \arctan, \dots)$ on
the individual measurements.
In programming languages such as C, this would involve looping over every element
to perform these operations.
Array computations generalize these element-wise operations on scalars to
vectors (1-d arrays), matrices (2-d arrays), and higher dimensional arrays (n-d
arrays).
For instance, if \code{x} is an \code{ndarray}, then \code{2 * x} produces
an \code{ndarray} whose values are twice those in \code{x}.
Similarly, if \code{x} and \code{y} have the same shape, then
\code{x * y} produces an \code{ndarray} whose values are the products of
the corresponding entries of \code{x} and \code{y}.
These \emph{vectorized} operations often produce one-liners that would tens
or hundreds of lines of code in languages such as C.

@@

\section{Recent technical improvements}

@@

\subsection{The Array Function Protocol}

@@

\subsection{Random}

@@

\section{Discussion}

>>>>

@@

\section*{History}

@@

\section*{Enabling factors}

@@

\section*{Hybrid development}

@@

% BIDS -- UCB
% https://www.moore.org/grant-detail?grantId=GBMF5447
% $645,020 in 2016
% https://sloan.org/grant-detail/8222
% $659,359 in 2017
% https://bids.berkeley.edu/news/bids-receives-sloan-foundation-grant-contribute-numpy-development
% http://doi.org/10.5281/zenodo.3585761
% http://doi.org/10.5281/zenodo.3585767
In 2017, NumPy received its first large grants totaling 1.3M USD from the
Gordon \& Betty Moore and the Alfred P. Sloan foundations.
These two grants focus on addressing the technical debt accrued over the years and
on setting in place standards and architecture to encourage more sustainable development.
% CZI -- NumFOCUS/QuanSight
% https://chanzuckerberg.com/eoss/proposals/strengthening-numpys-foundations-growing-beyond-code/
% NumPy and OpenBLAS received $195,000 in 2019
% https://labs.quansight.org/blog/2019/11/numpy-openblas-CZI-grant/
NumPy received a third grant for 195K USD from the Chan Zuckerberg
Initiative at the end of 2019.
This grant focuses on better serving NumPy's large number of beginning
to intermediate level users and on growing the community of NumPy
contributors.
Finally, the project receives around 10K USD annually from Tidelift, which is
typically used to fund documentation writers and website developers.

@@

\section*{NumPy arrays}

@@

The \code{ndarray} data structure stores regularly spaced elements of a single
data type in a contiguous block of memory; along with meta-data such as shape,
it allows for the efficient representation of $n$-dimensional arrays \cite{vanderwalt2011numpy}.
The need for such a data structure arises often in science.
For instance, they may store measurements from an experiment or simulation
at discrete time points.

Data analysis frequently involves operations $(+, -, \times, /, \log, \sin, \dots)$ on
the individual measurements.
In programming languages such as C or Java, this would involve looping over
every element to perform these operations.
Array computations generalize these element-wise operations on scalars to
vectors (1-d arrays), matrices (2-d arrays), and higher dimensional arrays (n-d
arrays).
For instance, if \code{x} is an \code{ndarray}, then \code{2 * x} produces
an \code{ndarray} whose values are twice those in \code{x}.
Similarly, if \code{x} and \code{y} have the same shape, then
\code{x * y} produces an \code{ndarray} whose values are the products of
the corresponding entries of \code{x} and \code{y}.
These \emph{vectorized} operations often produce one-liners that would be tens
or hundreds of lines of code in languages such as C.

@@

@@

\section*{Array function protocol}

@@

\section*{Random}

@@

\section*{Discussion}
